The likelihood surface

### 2017-04-21 shape/likelihood-surface.Rmd
more documentation changed update() to update_x() added a likelihood function factory

library(MASS)
library(numDeriv)
library(pander)
if (!file.exists(modelfile)) { stop("Can't find ", modelfile) }
model.config <- fill.default.config( read.config(modelfile) )
model.config$fixfn <- parse.fixfn( model.config$fixfn, model.config$fixfn.params )
if (!exists("gmfile")) {
    gmfile <- file.path(dirname(modelfile),gsub("%",opt$longwin,model.config$genmatrix,fixed=TRUE))
}
if (file.exists(gmfile)) {
    load(gmfile)
} else { 
    cat("Can't find generator matrix ", gmfile, " -- making a new one.\n")
    genmatrix <- do.call( makegenmatrix,
                    c( list(
                        patlen=opt$longwin, 
                        mutpats=model.config$mutpats, 
                        selpats=model.config$selpats, 
                        selfactors=model.config$selfactors, 
                        boundary=opt$boundary, 
                        bases=model.config$bases, 
                        fixfn=model.config$fixfn ),
                        model.config$fixfn.params
                    )
                )
}
genmatrix@x <- do.call( update, c( list(G=genmatrix, mutrates=model.config$mutrates, selcoef=model.config$selcoef), model.config$fixfn.params ) )
projmatrix <- collapsepatmatrix( ipatterns=rownames(genmatrix), leftwin=opt$leftwin, rightwin=opt$longwin-opt$shortwin-opt$leftwin, bases=genmatrix@bases )
```

@@

likfun <- function (sub.params){
    # params are: mutrates, selcoef, fixparams
    params[use.par] <- sub.params
    fparams <- params[seq( 1+nmuts(genmatrix)+nsel(genmatrix), length.out=length(fixparams(genmatrix)) )]
    names(fparams) <- fixparams(genmatrix)
    genmatrix@x <- do.call( update, c( list( G=genmatrix,mutrates=params[1:nmuts(genmatrix)],selcoef=params[seq(1+nmuts(genmatrix),length.out=nsel(genmatrix))]), as.list(fparams) ) )
    # this is collapsed transition matrix
    subtransmatrix <- computetransmatrix( genmatrix, projmatrix, tlen=1, time="fixed") # shape=params[length(params)], time="gamma" )
    # return POSITIVE log-likelihood
    ans <- sum( counts@counts * log(subtransmatrix) )
    if (!is.finite(ans)) { print(paste("Warning: non-finite likelihood with params:",paste(params,collapse=", "))) }
    return(ans)
}

>>>>

library(MASS)
library(numDeriv)
library(pander)
if (!file.exists(modelfile)) { stop("Can't find ", modelfile) }
model.config <- fill.default.config( read.config(modelfile) )
model.config$fixfn <- parse.fixfn( model.config$fixfn, model.config$fixfn.params )
if (!exists("gmfile")) {
    gmfile <- file.path(dirname(modelfile),gsub("%",opt$longwin,model.config$genmatrix,fixed=TRUE))
}
if (file.exists(gmfile)) {
    load(gmfile)
} else { 
    cat("Can't find generator matrix ", gmfile, " -- making a new one.\n")
    genmatrix <- do.call( makegenmatrix,
                    c( list(
                        patlen=opt$longwin, 
                        mutpats=model.config$mutpats, 
                        selpats=model.config$selpats, 
                        selfactors=model.config$selfactors, 
                        boundary=opt$boundary, 
                        bases=model.config$bases, 
                        fixfn=model.config$fixfn ),
                        model.config$fixfn.params
                    )
                )
}
genmatrix@x <- do.call( update_x, c( list(G=genmatrix, mutrates=model.config$mutrates, selcoef=model.config$selcoef), model.config$fixfn.params ) )
projmatrix <- collapsepatmatrix( ipatterns=rownames(genmatrix), leftwin=opt$leftwin, rightwin=opt$longwin-opt$shortwin-opt$leftwin, bases=genmatrix@bases )
```

@@

likfun <- function (sub.params){
    # params are: mutrates, selcoef, fixparams
    params[use.par] <- sub.params
    fparams <- params[seq( 1+nmuts(genmatrix)+nsel(genmatrix), length.out=length(fixparams(genmatrix)) )]
    names(fparams) <- fixparams(genmatrix)
    genmatrix@x <- do.call( update_x, c( list( G=genmatrix,mutrates=params[1:nmuts(genmatrix)],selcoef=params[seq(1+nmuts(genmatrix),length.out=nsel(genmatrix))]), as.list(fparams) ) )
    # this is collapsed transition matrix
    subtransmatrix <- computetransmatrix( genmatrix, projmatrix, tlen=1, time="fixed") # shape=params[length(params)], time="gamma" )
    # return POSITIVE log-likelihood
    ans <- sum( counts@counts * log(subtransmatrix) )
    if (!is.finite(ans)) { print(paste("Warning: non-finite likelihood with params:",paste(params,collapse=", "))) }
    return(ans)
}
